---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning = FALSE
)
```

# autovi

<!-- badges: start -->
<!-- badges: end -->

Buja et al. (2009) provides an inferential framework to assess whether residual plots indeed contain visual patterns inconsistent with the model assumptions. However, unlike conventional statistical tests that can be performed computationally in statistical software, the lineup protocol requires human evaluation of images. This characteristic makes it less suitable for large-scale applications, given the associated high labour costs and time requirements.

The `autovi` package aims to offer tools for automated visual inference of residual plots. Currently, it only supports diagnostic checks for classical normal linear regression models (CNLRM), as the underlying computer vision models are specifically trained for this model class. However, the `autovi` API is designed to be extensible to other model types. This means you can obtain predictions as long as you provide an appropriate method for generating null residuals through the `null_method` argument.

## Installation

You can install the development version of autovi like so:

``` r
remotes::install_github("TengMCing/autovi")
```

## Trained Keras Models

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(autovi)
```

All the available trained Keras models are listed in `list_keras_models()`. All the trained Keras models listed below will predict a visual signal strength for the visual patterns of the input residual plot. This visual signal strength is essentially an approximation of a Kullback–Leibler divergence based distance metric which quantifies the difference between the actual residual distribution and the reference residual distribution assumed under correct model specification. Details about the methodologies are provided in our draft paper https://github.com/TengMCing/auto_residual_reading/blob/master/paper/paper.pdf.

```{r}
list_keras_models() %>% pull(model_name)
```

Different Keras models are trained with residual plots obtained from different linear regression models violating various model assumptions. 

`vss_32`, `vss_64` and `vss_128` are trained with residual plots containing visual patterns indicating non-linearity and heteroskedasticity issues. The number in the model name represents the size of the input image. For instance, `32` means the input image has $32 \times 32$ pixels. 

`vss_phn_32`, `vss_phn_64` and `vss_phn_128` are also trained with residual plots containing visual patterns indicating non-linearity and heteroskedasticity, but the visual patterns are more complex. Additionally, residual plots with visual patterns of non-normality are also used for these models. 

## Examples

For this example, we will be using the `vss_phn_32` Keras model. The model can be obtained by using `get_keras_model()`.

```{r}
keras_model <- get_keras_model("vss_phn_32")
```

To illustrate the use of this package, we will define an incorrectly specified model omitting certain higher-order terms of $x$.

```{r}
set.seed(2024)
x <- rnorm(300)
y <- 1 + x + x^2 + x^3 + rnorm(300, sd = 3)
this_model <- lm(y ~ x)
```

The residual plot of the fitted model shows a "S" shape indicating a non-linearity issue.

```{r}
ggplot() +
  geom_point(aes(this_model$fitted.values, this_model$residuals)) +
  xlab("Fitted values") +
  ylab("Residuals")
```


To run diagnostics for this fitted model, we can initialize the checker using `auto_vi()` and have a look at the input residual plot using the `plot_resid()` method.

```{r}
checker <- auto_vi(fitted_model = this_model, keras_model = keras_model)
checker$plot_resid()
```

To predict the visual signal strength of this residual plot, simply use the `vss()` method.

```{r}
checker$vss()
```

Having the visual signal strength of the residual plot is usually insufficient to determine if the model is correctly specified. Thus, we need to evaluate some null residual plots for comparison.

The checker includes a default method to generate null residuals consistent with the null hypothesis that the fitted model is correctly specified, which is the `rotate_resid()` method implementing the residual rotation technique. This method is only suitable for CLNRM. 

```{r}
checker$rotate_resid() %>%
  checker$plot_resid()
```

To get predictions for null residual plots, one can use the `null_vss()` method. The `keep_null_data` and `keep_null_plot` tells the method whether to preserve the null residuals and null residuals plots in the result. For models that needs to use a null generating method other than `rotate_resid()`, the function can be provided via the `null_method` argument. The only parameter of the provided function should be `fitted_model`, which is the fitted model object. And it should return a data frame with two columns `.fitted` and `.resid` which are fitted values and null residuals respectively.

```{r}
checker$null_vss(20L, 
                 keep_null_data = FALSE, 
                 keep_null_plot = FALSE)
```

If we want to measure the variation of the visual signal strength of the residual plot, we can use the `boot_vss()` method to get bootstrapped visual signal strength. This method resamples the observations with replacement and refitt the regression model. Similarly, `keep_boot_data` and `keep_boot_plot` tells the method whether to preserve bootstrapped residuals and plots.


```{r}
checker$boot_vss(20L, 
                 keep_boot_data = FALSE, 
                 keep_boot_plot = FALSE)
```

To run a comprehensive check including the analysis of null residuals and bootstrapped residuals, use the `check()` method.

```{r}
checker$check(null_draws = 20L, boot_draws = 20L)
```

The check result is stored in the `check_result` attribute. If you print the `checker` object, it will show a brief summary of the result.

```{r}
checker
```

A summary plot can be drawn with the `summary_plot()` method. The solid line is the 95% sample quantile of the visual signal strength predicted for null residual plots. The dot line is the visual signal strength predicted for the original residual plot. The blue density curve indicates the distribution of visual signal strength predicted for null residual plots. And the red density curve indicates the distribution of visual signal strength predicted for bootstrapped residual plots.

The $p$-value is obtained by computing the ratio of null visual signal strength greater than or equal to observed visual signal strength. The red area along with the solid line indicates how often the fitted model would be considered incorrectly specified if the data can be repetitively drawn from the same data generating process.

For this example, the $H_0$ is rejected because of small $p$-value. It can also be observed that almost all the time the refitted models will be considered as incorrectly specified, so this is a clear rejection. Furthermore, since two density curves are very different from each other, it is very unlikely the original residuals are from the same distribution of null residuals.

```{r}
checker$summary_plot()
```

### Feature extraction

This package also enables the extraction of features from keras model for other purposes. To extract features from a specific layer of the keras model, one needs to provide the argument `extract_feature_from_layer`. The following code chunk extract features from the global pooling layer of the keras model.

```{r}
checker$vss(extract_feature_from_layer = "global_max_pooling2d") %>%
  names()
```

To check all the available layer names, one can list them with the `KERAS_WRAPPER` class.

```{r}
KERAS_WRAPPER$list_layer_name(keras_model)
```

While running the comprehensive check with the `check()` method, one can provide the `extract_feature_from_layer` argument to extract features.

```{r}
checker$check(null_draws = 20L, boot_draws = 20L, extract_feature_from_layer = "global_max_pooling2d")
```

The features are also stored in the `check_result` attribute.

```{r}
checker$check_result$observed
```

This package provide ways to conduct PCA on these features. The `feature_pca` method will combine the features obtained from the prediction of the observed residual plot, null residual plots, and the bootstrapped residual plots, then conduct PCA on them. The results are the rotated features. The variable `set` provides the set label of the observation, which can either be "observed", "null" or "boot".

```{r}
checker$feature_pca()
```

